The importance of measurement uncertainty in ecological management

### 2018-01-29 manuscripts/manuscript.Rmd
update sims

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7)
library(tidyverse)
library(gridExtra)
```

```{r optional_plot_fonts, message=FALSE}
library(ggplot2)
library(Cairo)
library(extrafont)
library(hrbrthemes)

extrafont::loadfonts()
knitr::opts_chunk$set( dev="cairo_pdf")
hrbrthemes::import_roboto_condensed() 
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
```

```{r plot_colors}
#palette <- ggthemes::ptol_pal()(6)
palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))

@@

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.  This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknolwedged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.  Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigour of optimization to resolve a long-standing paradox and suggest a more robust approach to management.

@@

A few subtlies arise in this seemlingly simple problem statement that we must address.  First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.

@@

To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),

@@

This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis which does not solve the time-dependent optimization problem above. In particular, this maximum sustainable yield (MSY) strategy will not be optimal whenever the stock is away from $B_{MSY}$.  @Clark1973 deomonstrated (assuming no uncertainty in measurement or stochasticity in population dynamics) that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*. We will return to this approach in a moment.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.  Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above. Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarise here.

@@

The first of these we shall refer to as Reed's Paradox for stochastic growth, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.  This suprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.  This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.

@@

Unforunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:

@@

Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than rejct it.  Others have also attempted to introduce measurement error, but always relying on various approximations that similarly alter the problem. The most common such assumption is to assume from the outset that the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.  In addition to resolving this paradox, we will demonstrate this direct solution also out-performs current more heuristic rules of thumb such as those currently used in fisheries management to account for uncertainty.

@@

- Utility function: $U(x_t,a_t)$, the utility recieved at time $t$ for taking action $a_t$, given that the system is in state $x_t$.  For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.  This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation. More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests. By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.

The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time. Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.  Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intellegence [. @Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.

@@

```{r sims, fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty. Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method. Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.  Environmental stochasticity is fixed a standard deviation of 0.15 in each panel.  Carrying capacity K normalized to 1, r = 0.75. Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material. \\label{sims}"}
sims %>%
  select(-value) %>%
  filter(method != "msy", sigma_g == "0.15") %>%
  mutate(method = fct_recode(method, 
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP" = "pomdp")) %>%
  group_by(time, method, sigma_m) %>%
  summarise(mean = mean(state), sd = sd(state)) %>% 
  mutate(sigma_m = as.character(sigma_m)) %>%
  ungroup() ->
fig1_df

fig1_df %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line(lwd=1) + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  coord_cartesian(ylim = c(0, 1)) + 
  theme(legend.position = "bottom") + 
  ylab("Mean biomass")
```

Figure \ref{sims} shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.  Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.  In each panel shown in Figure \ref{sims} stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.  The first panel of Figure \ref{sims} shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.  Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.

Figure \ref{sims} confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.  In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.  Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.  In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.

```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$. Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{{sims}}. The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error. The more cautious TAC is not economically optimal but largely unimpacted by increasing error, while POMDP attains consistently high economic yield despite the increasing uncertainty. \\label{econ}"}
sims %>%
  select(time, value, rep, method, sigma_m, sigma_g) %>%
  filter(sigma_g == "0.15", method != "msy") %>%
  group_by(rep, method, sigma_m, sigma_g) %>%
  summarise(npv = sum(value)) %>%  
  group_by(method, sigma_m, sigma_g) %>%
  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
  ungroup() %>%
  mutate(method = fct_recode(method,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP" = "pomdp")) %>%

  ggplot(aes(method, net_value, ymin=net_value-se, ymax=net_value+se, fill=method)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(size=.3,width=.2,
                position=position_dodge(.9)) +
  facet_wrap(~sigma_m, 
             labeller = label_bquote(sigma[m] == .(sigma_m))) +
  theme(legend.position = "none") + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  ylab("expected net present value")
```

This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.  Figure \ref{econ} shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.  As before, environmental stochasticity is set at $\sigma_g = 0.15$.  In the absense of measurement uncertainty, constant escapement (CE) is optimal (as per @Reed1979), while the heuristic caution built into the total allowable catch (TAC) strategy (at 80% MSY) results in a sub-optimal economic yield.  The presense of stochasticity in recruitment ($\sigma_g = 0.15$), also contributes to the reduction in yield under TAC.  Though CE is quite robust to the environmental stochasticity level, this strategy proves very sensitive to increasing measurement error, showing sharp declines in economic value, falling below the TAC economic value at $\sigma_m = 0.15$.    Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure \ref{sims}), these have little impact on the economic value due to the discount rate.  A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.  Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.  In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absense of measurement error) despite the increasingly uncertain measurements.  This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims}) is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.

>>>>

```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7,  dev="cairo_pdf")
library(tidyverse)
library(gridExtra)
```

```{r optional_plot_fonts, message=FALSE}
library(ggplot2)
library(Cairo)
library(extrafont)
library(hrbrthemes)
extrafont::loadfonts()
knitr::opts_chunk$set()
hrbrthemes::import_roboto_condensed() 
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
palette <- c("#D9661F", "#3B7EA1",  "#6C3302", 
             "#FDB515", "#00B0DA",  "#CFDD45")
colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: high prior",
                      "biomass",
                      "catch"))
scale_colour_discrete <- function(...) scale_colour_manual(..., values=colors)
scale_fill_discrete <- function(...) scale_fill_manual(..., values=colors)
```

@@

Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierarchical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and under-served by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these existing approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.  This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknowledged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.  Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigor of optimization to resolve a long-standing paradox and suggest a more robust approach to management.

@@

A few subtleties arise in this seemingly simple problem statement that we must address.  First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment.

@@

To facilitate tractability and interpretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),

@@

This *constant yield* (constant mortality) solution thus corresponds to an equilibrium analysis which does not solve the time-dependent optimization problem above. In particular, this maximum sustainable yield (MSY) strategy will not be optimal whenever the stock is away from $B_{MSY}$.  @Clark1973 demonstrated (assuming no uncertainty in measurement or stochasticity in population dynamics) that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*. We will return to this approach in a moment.

@@

Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC, ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Over-fishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.

@@

So far, attempts to provide a more formal basis for managing uncertainty than the heuristic adjustment of catch limits described above have largely foundered in a series of of paradoxes.  Without uncertainty, the theoretical policies are quite intuitive: The result derived by @Clark1973 for an optimal dynamic strategy to replace the equilibrium solution of MSY can be summarized as: If the stock is most productive at $B_{MSY}$, then obtain $B_{MSY}$ as quickly as possible. Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, harvests greater than $H_{MSY}$ will be needed to bring the stock back to $B_{MSY}$.  (This intuition must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.  Intuitively, any uncertainty, either in the population growth model (stochasticity), or measurement error, should seem to justify harvesting fewer fish, as is built into the TAC limits discussed above. Yet formal treatments of these issues have so far required approximations, and these approximations have driven paradoxical conclusions which we briefly summarize here.

@@

The first of these we shall refer to as Reed's Paradox for stochastic growth, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.  This surprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.  This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.

@@

Unfortunately for @Clark1986, this "simplifying assumption" serves to squash most of the measurement error, and their results instead only deepened the paradox, finding policies that become even *less* cautious as uncertainty increases:

@@

Despite these notes of caution, both @Clark1986 and @Sethi2005 ultimately attempt to rationalize this counter-intuitive conclusion rather than reject it.  Others have also attempted to introduce measurement error, but always relying on various approximations that similarly alter the problem. The most common such assumption is to assume from the outset that the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.  In addition to resolving this paradox, we will demonstrate this direct solution also out-performs current more heuristic rules of thumb such as those currently used in fisheries management to account for uncertainty.

@@

- Utility function: $U(x_t,a_t)$, the utility received at time $t$ for taking action $a_t$, given that the system is in state $x_t$.  For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.  This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation. More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests. By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.

The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time. Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.  Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  While the solution method for MDP problems, stochastic dynamic programming (SDP) is well known and readily implemented in ecological problems [e.g @Mangel1988; @Marescot2013], the POMDP problem cannot be efficiently solved using SDP. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intelligence [. @Smallwood1973; @Sondik1978; @Pineau2003; @Kaelbling1998], and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries management, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.

@@

```{r sims, fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty. Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method. Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.  Environmental stochasticity is fixed a standard deviation of 0.15 in each panel.  Carrying capacity K normalized to 1, r = 0.75. Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material. \\label{sims}"}
sims %>%
  select(-value) %>%
  filter(method != "MSY", sigma_g == "0.15") %>%
  group_by(time, method, sigma_m) %>%
  summarise(mean = mean(state), sd = sd(state)) %>% 
  mutate(sigma_m = as.character(sigma_m)) %>%
  ungroup() ->
fig1_df

fig1_df %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line(lwd=1) + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
  coord_cartesian(ylim = c(0, 1)) + 
  theme(legend.position = "bottom") + 
  ylab("Mean biomass")
```

Figure \ref{sims} shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.  Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.  In each panel shown in Figure \ref{sims} stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.  The first panel of Figure \ref{sims} shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.  Without measurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.

Figure \ref{sims} confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.  In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.  Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC strategy rather than the CE strategy that lead to the highest over-fishing rate.  In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.

```{r econ, fig.width=7, fig.height=3.5, fig.cap="Expected net present value of the fishery across strategies under increasing levels of measurement error $\\sigma_m$. Net present value is th sum of all future harvests discounted over time and averaged across 100 replicates simulations under each strategy, as shown in Figure \\ref{{sims}}. The value under constant escapement (CE) is optimal when measurements are perfect but decreases rapidly with increasing measurement error. The more cautious TAC is not economically optimal but largely unimpacted by increasing error, while POMDP attains consistently high economic yield despite the increasing uncertainty. \\label{econ}"}
sims %>%
  select(time, value, rep, method, sigma_m, sigma_g) %>%
  filter(sigma_g == "0.15", method != "MSY") %>%
  group_by(rep, method, sigma_m, sigma_g) %>%
  summarise(npv = sum(value)) %>%  
  group_by(method, sigma_m, sigma_g) %>%
  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
  ungroup() %>%

  ggplot(aes(method, net_value, ymin=net_value-se, ymax=net_value+se, fill=method)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(size=.3,width=.2,
                position=position_dodge(.9)) +
  facet_wrap(~sigma_m, 
             labeller = label_bquote(sigma[m] == .(sigma_m))) +
  theme(legend.position = "none") + 
  ylab("expected net present value")
```

This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also borne out in terms of economic performance.  Figure \ref{econ} shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.  As before, environmental stochasticity is set at $\sigma_g = 0.15$.  In the absence of measurement uncertainty, constant escapement (CE) is optimal (as per @Reed1979), while the heuristic caution built into the total allowable catch (TAC) strategy (at 80% MSY) results in a sub-optimal economic yield.  The presence of stochasticity in recruitment ($\sigma_g = 0.15$), also contributes to the reduction in yield under TAC.  Though CE is quite robust to the environmental stochasticity level, this strategy proves very sensitive to increasing measurement error, showing sharp declines in economic value, falling below the TAC economic value at $\sigma_m = 0.15$.    Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure \ref{sims}), these have little impact on the economic value due to the discount rate.  A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.  Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.  In contrast to the declining economic performance of CE and the consistently sub-optimal economic yield for TAC, the POMDP strategy continues to generate economic yields at approximately the optimal level (as attained by CE in the absence of measurement error) despite the increasingly uncertain measurements.  This demonstrates that the reduced risk of stock collapse and higher average stock biomass attained by the POMDP strategy in the presence of high measurement error, Figure \ref{sims}) is not the result of a trivial reduction in harvesting under all circumstances, but rather, evidence of a more nuanced strategy that manages to account for the uncertainty in measurement while maintaining a reasonable harvest.
